### 2021-03-10 eda.Rmd
Loaded data and libraries, and displayed first several rows

---
title: "R Notebook"
output: html_notebook
---

```{r include=FALSE}
knitr::opts_chunk$set('echo'=FALSE)

library(dplyr)
library(ggplot2)

load('streetrx.rdata')
```

```{r}
head(streetrx, 15)
```