### 2021-03-10 eda.Rmd
Beginning of analysis

---
title: "R Notebook"
output: html_notebook
---

# Todo

- Check for missing data across columns (including response), as well as alternate ways data might be NA
- Perform EDA to sanity check values, check types, etc.

# Changes to data

- Filtered data such that `api_temp = oxymorphone`
- Dropped `country` and `api_temp`
- Removed observations with missing `ppm` (14)

```{r include=FALSE}
knitr::opts_chunk$set('echo'=FALSE)

library(dplyr)
library(ggplot2)
library(lubridate)

load('streetrx.rdata')
streetrx <- streetrx %>% filter(api_temp == 'oxymorphone')

# Only 1 country (USA), and we're only working with 
drop_cols <- c('country', 'api_temp')
streetrx <- streetrx %>% select(-any_of(drop_cols))
```

```{r}
set.seed(42)
slice_sample(streetrx, n=20)
```

# Initial observations (brain dump)

- At least one of the response variables (`ppm`) is missing - we'll need to remove these observations
- `yq_pdate` and `price_date` contain redundant information - possibly best to break these 2 columns apart into quarter, day, month, and year columns. Likely day won't be important after accounting for month and year - check a random set of same month/year pairings and see if there's a trend?
- `city` frequently missing
- `state` contains 59 levels: will need to look into this
- `country` column not needed (only working with USA)
- `source` looks frequently missing and it's not clear how many levels there are: look into
- `api_temp`: Not needed (we're only working with oxymorphone)
- `mgstr`: Will need to validate this - shouldn't be negative or 0 (check histogram)
- `bulk_purchase`: Remove the text
- `primary_reason`: Looks missing frequently

# Checking Columns One-byone

The big one. First, let's take a look at which fields have missing values and how frequently they occur. First, the response variable

```{r}
sum(is.na(streetrx$ppm))
```

We'll need to drop these, because a missing response provides no assistance to our model:

```{r}
streetrx <- streetrx %>% filter(!is.na(ppm))
```

Any other weird values?

```{r}
ggplot(streetrx, aes(x=ppm)) + geom_histogram(bins = 50)
```

There are some very extreme values here:

```{r}
streetrx %>% filter(ppm > 10)
```

I'm certainly not an expert on drugs. But based on the previous prices, these are likely erroneously recorded, or not serious entries. Keep an eye on and consider dropping - the corresponding points could be influential