### 2021-03-14 cjuracek_model_building.Rmd
Added model fitting with lme4 and brms

---
title: "Cole Juracek - Model Building"
output: html_notebook
---

```{r}
library(lme4)
library(lmerTest)
library(brms)
library(dplyr)

load('mod_streetrx.Rda')
streetrx$year <- as.numeric(as.character(streetrx$year))

# Need to relabel single observation
streetrx$source <- recode_factor(streetrx$source, 'STREET PRICE' = 'Heard it')

# Similar reason
levels(streetrx$Primary_Reason)[levels(streetrx$Primary_Reason) == ''] <- '0 Reporter did not answer the question'
```

# Simple models

No interactions, random intercept on state (no other random slopes), no Bayesian specification:

Predicting: log_ppm

Which variables to we want to include?

- State (as a random effect)
- USA_region
- source
- mgstr
- bulk_purchase
- Primary reason
- year

```{r}
#model_1 <- lmer(log_ppm ~ (1 | state) + (source + mgstr + bulk_purchase + year)^2, data=streetrx)
model_1 <- lmer(log_ppm ~ (1 | state) + source + mgstr + bulk_purchase + year, data=streetrx)
BIC(model_1)
```

Add new random effects:

```{r}
model_2 <- lmer(log_ppm ~ (1 | state) + USA_region + source + mgstr + bulk_purchase +
                Primary_Reason + year, data=streetrx)
#BIC(model_2)
step(model_2)
```

```{r}
model_bms <- brm(data = streetrx, family = gaussian,
    log_ppm ~ (1 | state) + source + mgstr + bulk_purchase + year,
    prior = c(prior(normal(0, 10), class = Intercept),
              prior(normal(0, 10), class = b),
              prior(cauchy(0, 1), class = sd),
              prior(cauchy(0, 1), class = sigma)),
    iter = 100, warmup = 10, chains = 2, cores = 2,
    seed = 13)
```