### 2021-03-15 hm-cs1-work-cleaned-3_14.Rmd
Rmd file including EDA, model selection, diagnostics...

---
title: "hm-cs1"
output: html_document
---

```{r message=FALSE}
library(tidyverse)
library(lme4)
```

```{r}
load("streetrx.RData")
```

```{r}
df <- streetrx %>% filter(api_temp=="oxymorphone")
```

# Cleaning individual variables

* Response: ppm
* Grouping variables to consider: city, state, USA_region
* Non-group predictors to consider: Continuous: yq_pdate, mgstr, ; Categorical: source, bulk_purchase, Primary_Reason
* Can drop: price_date, country (all USA), api_temp, form_temp

```{r}
# cleaning state:

# table(df$state)

# remove levels corresponding to states with 0 observations in df
df$state <- droplevels(df$state)
# table(df$state)

# drop North Dakota - one observation
df <- df %>% filter(state != "North Dakota")
df$state <- droplevels(df$state)
# table(df$state)

# replace state "USA" with "Unknown"
df$state <- recode_factor(df$state, "USA" = "Unknown")
# table(df$state)

```

```{r}
# cleaning region:

# # looks fine:
# table(df$USA_region)
# 
# # confirms that the unknown state observations are same as those with unknown region
# table(df$USA_region, df$state)
```

Change date to numeric (easier for interpretation?):

```{r}
# hist(as.numeric(df$price_date))

# this sets 
df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 10957 # sets 1/1/2000 as zero

tempdf <- df %>% select(date_num, price_date)

# this filters out all data before the 1/1/2000, which probably shouldn't be included
df <- df %>% filter(date_num >= 0)

hist(df$date_num)
```

```{r}
hist(df$mgstr)
# fine as is?
```

```{r}
# recode sources

# remove unused levels
df$source <- droplevels(df$source)

source_df <- data.frame(table(df$source))

# levels(df$source)
levels(df$source)[1] <- "Unknown"

source_df <- data.frame(table(df$source))

df$source <- recode_factor(df$source, "Unknown" = "Unknown",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL

source_df_2 <- data.frame(table(df$source))

print(source_df)
print(source_df_2)
```

```{r}
table(df$bulk_purchase, useNA = "always")
# looks fine
```

```{r}

# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)

reason_df <- data.frame(table(df$Primary_Reason))

# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Other or unknown"

reason_df <- data.frame(table(df$Primary_Reason))

df$Primary_Reason <- recode_factor(df$Primary_Reason, "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                           .default = "Other or unknown")

reason_df_2 <- data.frame(table(df$Primary_Reason))

print(reason_df)
print(reason_df_2)

```

```{r}
# log transform for ppm looks good
hist(log(df$ppm))

# transform
df$log_ppm <- log(df$ppm)
```

## More EDA

Thinking about what to group by...
State best? Regions kind of similar, cities have many small sample sizes...

```{r}
ggplot(df, aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state", x="state",y="Log PPM")

ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")

ggplot(df[df$state=="Florida",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="city",y="Log PPM")

ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (CA)", x="city",y="Log PPM")
```

Closer look at by-state values of price

```{r}
ggplot(df[df$USA_region == "Northeast" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - Northeast", x="state",y="Log PPM")

ggplot(df[df$USA_region == "West" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - West", x="state",y="Log PPM")

ggplot(df[df$USA_region == "South" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - South", x="state",y="Log PPM")

ggplot(df[df$USA_region == "Midwest" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - Midwest", x="state",y="Log PPM")

ggplot(df[df$USA_region == "Other/Unknown" ,], aes(x=reorder(state, state, length), y=log_ppm, fill=state)) +
geom_boxplot() +
labs(title="log PPM by state - Other/Unknown", x="state",y="Log PPM")
```

Looking more at cities...

```{r}
# some states have few observations for even most populous cities... 
temp <- df$city[df$state=="Massachusetts"]
length(temp)

# look at the big cities overall 
city_temp <- df %>% group_by(city) %>% summarize(city_count = n()) %>% arrange(-city_count)
city_temp[1:20, ]
hist(city_temp$city_count[-1])
View(city_temp)

# look at the big states  
state_temp <- df %>% group_by(state) %>% summarize(state_count = n()) %>% arrange(-state_count)
state_temp[1:20, ]
hist(state_temp$state_count[-1])
View(state_temp)
```

Use random slopes (by region)?
Looks like no for continuous variables, maybe for categorical ones.

```{r}
ggplot(data = df, aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = source, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ USA_region)

ggplot(data = df, aes(x = Primary_Reason, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ USA_region)
```

Random slopes by state? Sampling a few states at a time for visibility.

For continuous variables looks like no, maybe for categorical variables.

```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Massachusetts" | df$state=="Connecticut" | df$state=="New York",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="California" | df$state=="Oregon" | df$state=="Arizona",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Michigan" | df$state=="Wisconsin" | df$state=="Kansas",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

```{r}
# random slopes for mgstr?

ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Massachusetts" | df$state=="Connecticut" | df$state=="New York",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="California" | df$state=="Oregon" | df$state=="Arizona",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)

ggplot(data = df[df$state=="Michigan" | df$state=="Wisconsin" | df$state=="Kansas",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

```{r}
# categorical variables -- looking at South only
ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)

ggplot(data = df[df$USA_region == "South",], aes(x = Primary_Reason, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)

ggplot(data = df[df$USA_region == "South",], aes(x = source, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

Modeling:

Include random intercepts for state, but should region also be included as fixed effect? -- No.
Are grouping by region or by city better than by state? -- No.

```{r}
mod1 <- lmer(data=df, log_ppm ~ (1 | state)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2 <- lmer(data=df, log_ppm ~ (1 | state) + USA_region + date_num + mgstr + bulk_purchase + Primary_Reason + source)
anova(mod1, mod2)
anova(mod1_2, mod1)
# confirsms we may be able to leave region out
# all other predictors in mod1 significant

mod_s <- lmer(data=df, log_ppm ~ (1 | state)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod_r <- lmer(data=df, log_ppm ~ (1 | USA_region)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod_c <- lmer(data=df, log_ppm ~ (1 | city)  + date_num + mgstr + bulk_purchase + Primary_Reason + source)

BIC(mod_s) # state grouping is lowest - confirms our suspicion that state was the best grouping variable
BIC(mod_r)
BIC(mod_c)
```

Try random slopes for each categorical variable. 
Do they give better model? -- No.

```{r}
# none of the random slopes for the categorical variables seem to add to the model 
# all predictors significant in mod1 so keep all 
# mod1 is final model

mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2_bulk <- lmer(data=df, log_ppm ~ (bulk_purchase| state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2_reason <- lmer(data=df, log_ppm ~ (Primary_Reason| state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
mod2_source <- lmer(data=df, log_ppm ~ (source| state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)

anova(mod1, mod2_bulk)
anova(mod1, mod2_reason)
anova(mod1, mod2_source)
```

Diagnostics of the (potential) final model: 
Good sign -- No states are outliers.

```{r}
library(influence.ME)
mod1_inf <- influence(mod1,"state")
print(2/sqrt(length(unique(df$state))))
# dfbetas(mod1_inf)
cooks.distance(mod1_inf,sort=TRUE)

# no states have values above the cutoff
```