### 2021-03-20 STA 610 Case Study 1 Report_Team 4.Rmd
Small formatting changes

---
title: "STA610 Case Study 1 Team 4 Report"
author: 
  - Cole Juracek (Coordinator) 
  - Lauren Palazzo (Programmer) 
  - Lingyu Zhou (Checker) 
  - Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
if (!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, lme4, gridExtra, grid, ggplot2, lattice, redres)
#devtools::install_github("goodekat/redres")
```

```{r, echo = FALSE}
load("streetrx.RData")
df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
```

```{r, echo = FALSE}
# cleaning state:

# table(df$state)

# remove levels corresponding to states with 0 observations in df
df$state <- droplevels(df$state)
# table(df$state)

# drop North Dakota - one observation
df <- df %>% filter(state != "North Dakota")
df$state <- droplevels(df$state)
# table(df$state)

# replace state "USA" with "Unknown"
df$state <- recode_factor(df$state, "USA" = "Unknown")
# table(df$state)

# log transform ppm
df$log_ppm <- log(df$ppm)
```

## EDA
#### PPM

```{r, fig.height=3, fig.align='center'}
hist_ppm <- ggplot(df, aes(x=ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="red")

hist_log_ppm <- ggplot(df, aes(x=log_ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="blue")

grid.arrange(hist_ppm, hist_log_ppm, ncol=2)
```
The histogram of the original ppm is extremely right skewed, while the distribution of the log(ppm) looks roughly normal. Thus, we will use the log(ppm) as the dependent variable in our model.

```{r, echo = FALSE}
# Change date to numeric (easier for interpretation)

# this sets 
df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 10957 # sets 1/1/2000 as zero

tempdf <- df %>% select(date_num, price_date)

# this filters out all data before the 1/1/2000, which probably shouldn't be included
df <- df %>% filter(date_num >= 0)

```

#### Sources

We recoded the categorical variable "sources" into the following 4 levels.
```{r, echo = FALSE}
# recode sources

# remove unused levels
df$source <- droplevels(df$source)

source_df <- data.frame(table(df$source))

# levels(df$source)
levels(df$source)[1] <- "Unknown"

source_df <- data.frame(table(df$source))

df$source <- recode_factor(df$source, "Unknown" = "Unknown",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL

source_df_2 <- data.frame(table(df$source))

knitr::kable(source_df_2, col.names = c("Level", "Freq"))
```

```{r}
#table(df$bulk_purchase, useNA = "always")
# looks fine
```

#### Primary_Reason

We recoded the categorical variable "Primary_Reason" into the following 3 levels.
```{r}
# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)

reason_df <- data.frame(table(df$Primary_Reason))

# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Other or unknown"

reason_df <- data.frame(table(df$Primary_Reason))

df$Primary_Reason <- recode_factor(df$Primary_Reason, "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                           .default = "Other or unknown")

reason_df_2 <- data.frame(table(df$Primary_Reason))

#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

#### State
```{r, fig.align='center', fig.height=6}
ggplot(df, aes(x=log_ppm, y=reorder(state, state, length), fill=state)) +
geom_boxplot() +
labs(title="state", x="log PPM by state",y="Log PPM")
```
The boxplot of log PPM by state is arranged in the order of increasing sample size from the bottom to the top.

#### Region

```{r, fig.align='center'}
ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")
```

#### City
```{r, fig.align='center'}
fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")

ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (CA)", x="Log PPM",y="city")

grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
```

#### mgstr
```{r, fig.align='center'}
# random slopes for mgstr?

ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = mgstr, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

#### bulk_purchase
```{r, fig.align='center'}
# categorical variables -- looking at South only
ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

#### date_num
```{r, fig.align='center'}
# categorical variables -- looking at South only
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

## Model

Per the research question of interest, we are interested in exploring factors related to price per mg of oxymorphone. We are also specifically asked to control for location and how drug price changes with it. Cities seemed to be too noisy to use as a location predictor. On the flip side, region did not seem to be associated with a significant change in drug prices. Therefore, state provides the best granularity of accounting for location in this instance.

Linear regression models are usually a good standard approach to this, providing clear interpretations of how various predictors are associated with drug price. As seen in the EDA, however, many states have excessively small sample sizes. This produces corresponding $\hat{\beta}_{OLS}$ estimates with large variance. We would like to be able to *partially pool* information across states, letting us borrow information to help stabilize state estimates with small sample sizes.

This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.

#### Model Specification

$$y_{ij} = \beta_{0,j} + \beta_{1} date_num + \beta_{2} mgstr + \beta_{3} bulk_purchase + \beta_{4} Primary_Reason + \beta_{5} source + \epsilon_{ij}$$
$$
\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
$$
$$
\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
$$

* i stands for individuals and j stands for each state. 
* $\beta_{0,j}$: Random intercept by state
* $\beta_{1}$: Fixed effect for date_num, the date variable
* $\beta_{2}$: Fixed effect for mgstr, the dosage strength of reported drug in milligrams.
* $\beta_{3}$: Fixed effect for bulk_purchase is a binary variable that is represented by an indicator function I(bulk_purchase = Yes). 
* $\beta_{4}$: Fixed effect for the categorical variable Primary_Reason which contains three categories:     - Other or unknown
    - Self-treat
    - Enjoymen
* $\beta_{5}$ measures the fixed effect for the categorical variable source

Model Selection discussions here.

```{r}
mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
```

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
# resqq <- plot_resqq(mod1)
# ranefci <- plot_ranef(mod1)
# grid.arrange(resqq, ranefci, ncol = 2)
```

Model diagnostics discussions here.

## Results
#### Fixed Effects
```{r, message=FALSE}
#knitr::kable(fixef(mod1))
#confint(mod1)[3:11,]

# extract coefficients
coefs <- data.frame(coef(summary(mod1)))
# use normal distribution to approximate p-value
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
knitr::kable(coefs, col.names = c("Estimate", "Std.Error", "t-value", "P-value"))
```

* Bulk_purchase: “Not bulk purchase” is the reference group.
Controlling for other variables, being bulk purchase increases the price per mg by exp(-0.16) = 0.8521. 
* Primary_Reason: “Primary_Reason: Other or Unknown” is the reference group. 
Controlling for other variables, taking oxymorphone with the primary reason of self-treating tends to receive a exp(0.05) = 1.0513 times higher price per mg than people whose primary reason is “other or unknown”.
Controlling for other variables, taking oxymorphone with the primary reason of enjoyment tends to receive a exp(0.02) = 1.0202 times higher price per mg than people whose primary reason is “other or unknown”.
* Source: “Source: Unknown” is the reference group. 
Controlling for other variables, purchasing oxymorphone with Internet source tends to receive a exp(0.02) = 1.0202 times higher price per mg than people with unknown drug source.
Controlling for other variables, purchasing oxymorphone with personal source tends to receive a exp(0.07) = 1.0725 times higher price per mg than people with unknown drug source.
Controlling for other variables, purchasing oxymorphone with “heard it” source tends to receive a exp(0.24) = 1.2712 times higher price per mg than people with unknown drug source.

#### Random Effects
```{r}
#VarCorr(mod1)
Variance <- c(0.04748, 0.96369)
Std.Dev <- c(0.2179, 0.9817)
Groups <- c("state (Intercept)", "Residual")

REvar <- cbind(Groups, Variance, Std.Dev)

knitr::kable(REvar, col.names = c("Group", "Variance", "Std.Dev"))

```
The within-state variance $\sigma^2$ is 0.96369, which is much larger than the across-state variance $\tau^2$, 0.04748.

```{r, fig.align='center', fig.height=8}
dotplot(ranef(mod1, condVar=TRUE))$state
```
From the result of the random effects, we can see that the random intercept for Tennessee is the highest and it is significantly different from zero. The random intercept for Nevada is the lowest and this intercept is also significant.

## Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```