### 2021-03-21 final_report.Rmd
Added my section for model selection and specification

---
title: "STA610 Case Study 1 Team 4 Report"
author:
- Cole Juracek (Coordinator)
- Lauren Palazzo (Programmer)
- Lingyu Zhou (Checker)
- Fan Zhu (Presenter)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)
# devtools::install_github("goodekat/redres")
library(redres)
library(stringr)
library(influence.ME)
```

```{r, echo = FALSE}
load("~/Documents/Spring 2021/STA 610 - H Models/cs1/streetrx.RData")
df <- streetrx %>% filter(api_temp=="oxymorphone", !is.na(ppm))
```

# Introduction

This study is an exploration of the relationship between the sale price per milligram of diverted pharmaceutical substances and other factors related to the sale, as reflected in self-reported data from the website StreetRx. StreetRx maintains a partnership with the Researched Abuse, Diversion, and Addiction-Related Surveillance System (RADARS), which aims for these data to inform public health measures such as health policy related to illicit drugs.

[insert something about our specific drug here]

Among the data available to us are location data pertaining to the city, state, and United States region of the sale. These variables present a natural grouping structure for the price data, which suggest that we employ a hierarchical model in our analysis. The benefits of a hierarchical model include the ability to "borrow information" from the data set as a whole to stabilize estimates of price for groups (e.g., states) with relatively small sample size, and to establish a estimated distribution of how price varies with respect to groups, which would in theory allow generalization to new groups not included in the original analysis.

# Exploratory data analysis and data cleaning

## Data cleaning

### Group variables

We performed an initial examination of two grouping variables, `state` and `USA_region`. The state North Dakota had only one observation, and as such was not considered a well-defined "group" for which a variance could be estimated; thus this state's observation was removed. States labeled simply as "USA" were recoded to "Unknown". Region was also sometimes labeled as uknown or other; these observations coincided with the unknown states (see appendix Table X). (The variable `city` as a grouping variable is addressed later in the report.)

```{r, echo = FALSE}
# cleaning state:
# table(df$state)
# remove levels corresponding to states with 0 observations in df
df$state <- droplevels(df$state)
# table(df$state)
# drop North Dakota - one observation
df <- df %>% filter(state != "North Dakota")
df$state <- droplevels(df$state)
# table(df$state)
# replace state "USA" with "Unknown"
df$state <- recode_factor(df$state, "USA" = "Unknown")
# table(df$state)
### PUT IN APPENDIX: the state-by-region table of coutns below
# table(df$state, df$USA_region)
```

### Date variable

We considered two different ways to engineer the date-of-sale feature. One intuitive and commonly-used approach would be to recode this as a continuous variables counting days from some start date, the intent of which would be to capture the influence of inflation for this price data that spans years. Another approach would be to consider whether there might be an effect of seasonality on drug prices.

For recoding date into a continuous variable, `date_num`, we filter out observations dated prior to the year 2010, as that is the year that the StreetRx web site was founded. For the seasonality variable, `date_quarter`, we extract the "quarter" of the year (numbered 1 through 4) as its own variable. Below we see the distributions of these variables:

```{r, echo = FALSE}
# Change date to numeric (easier for interpretation)
# this sets 
df$date_num <- as.numeric(as.Date(df$price_date, "%m/%d/%y")) - 14621 # sets 1/12/2010 (first post-2010 date) as zero
tempdf <- df %>% select(date_num, price_date)
# this filters out all data before the 1/1/2000, which probably shouldn't be included
df <- df %>% filter(date_num >= 0)
```

```{r}
df$date_quarter <- as.numeric(substr(df$yq_pdate, 5, 5))
```

```{r}
hist(df$date_num)
knitr::kable(table(df$date_quarter))
```

### Sources variable

As the variable `source` that describes the source of the price information is a user-input field, it is an inherently "messy" variable that calls for some recoding in order to make the levels of the variable more meaningful and interpretable, and to avoid overfitting. In the case of `source`, some user input is simply blank, while a small number of other entries indicate in some other way that the source is unknown or unclear. Since it is quite uncommon in the data for a source to be entered yet indicated as unknown, we decided to drop these unkown-source observations as being indicative of poor data quality. (If the user cannot identify their source of price information, why should we trust their price data?) We decided to distinguish between "Internet pharmacy" sources and other internet sources, since there may be a distinction between a source that is selling a drug versus perhaps just discussing a drug price in a forum, which may yield less legitimate information. Ultimately, we recoded the categorical variable "sources" into the following levels.

```{r, echo = FALSE}
# recode sources 
# remove unused levels
df$source <- droplevels(df$source)
source_df <- data.frame(table(df$source))
# levels(df$source)
levels(df$source)[1] <- "Blank"
source_df <- data.frame(table(df$source))
df$source <- recode_factor(df$source, "Blank" = "Blank",
                           "Internet" = "Internet",
                           "Internet Pharmacy" = "Internet Pharmacy",
                           "Personal" = "Personal",
                           "Heard it" = "Heard it",
                           "Idk" = "Unknown",
                           "w" = "Unknown",
                           "STREET PRICE" = "Heard it",
                           .default = "Internet") # everything else is a URL
source_df_2 <- data.frame(table(df$source))
# drop unknowns
df <- df %>% filter(df$source != "Unknown")
df$source <- droplevels(df$source)
knitr::kable(table(df$source), col.names = c("Level", "Freq"))
```

### Bulk purchase variable

Our visualization of the bulk purchase variable indicates nothing unexpected that calls for data cleaning.

```{r}
#table(df$bulk_purchase, useNA = "always")
# looks fine
```

### Purchase reason variable

`Primary_Reason` for purchase is another user-input variable that calls for recoding. Similarly to the `source` variable, we retain the information about which entries were blank. We also try to distinguish this from a different response level that is not blank, yet indicates the user did not provide a reason. We further still distinguished between these responses and a non-blank response of refusal to answer. We strove to avoid making presumptions of no difference between these similar-seeming response types, as they may indicate variations in data processing that hold some predictive power that we are not yet aware of. We ultimately recoded the categorical variable `Primary_Reason` into the following levels.

```{r}
# remove unused levels
df$Primary_Reason <- droplevels(df$Primary_Reason)
reason_df <- data.frame(table(df$Primary_Reason))
# levels(df$Primary_Reason)
levels(df$Primary_Reason)[1] <- "Blank"
reason_df <- data.frame(table(df$Primary_Reason))
df$Primary_Reason <- recode_factor(df$Primary_Reason, "Blank" = "Blank",
                                   "0 Reporter did not answer the question" = "Left unanswered",
                                   "Other or unknown" = "Other or unknown",
                                   "9 To self-treat my pain" = "Self-treat",
                                   "3 To prevent or treat withdrawal" = "Self-treat",
                                   "10 To treat a medical condition other than pain" = "Self-treat",
                                   "4 For enjoyment/to get high" = "Enjoyment",
                                   "5 To resell" = "Resell",
                                   "8 Prefer not to answer" = "Refuse to answer",
                           .default = "Other or unknown")
reason_df_2 <- data.frame(table(df$Primary_Reason))
#print(reason_df)
knitr::kable(reason_df_2, col.names = c("Level", "Freq"))
```

### Drug strength variable

We renamed this variable `dose_per_mg` to make its meaning clearer to the reader. This is a continuous variable that nevertheless tends to take on only a few values, as depicted in the histogram. We considered a recoding of this variable into levels, but decided against it as the dose per milligram technically may take on any value and the generalizability of our model may be improved if we are able to consider dose levels that are in a realistic range yet not present in our current data.

```{r}
hist(df$mgstr)
# rename variable for ease of interpretability
df <- df %>% rename(dose_per_mg = mgstr)
```

### Response variable: Price per milligram

We are considering fitting a hierarchical model with a response variable that has normally distributed errors, and so we would like our response variable, here `ppm`, or price per milligram, to be close to normally-distributed. The histogram of the `ppm` is extremely right skewed (not unexpected for data related to financial measures such as price, salary, earnings, etc.), while the distribution of the log(`ppm`) looks roughly normal. Thus, we will use `log_ppm` as the dependent variable in our model.

```{r, fig.height=3, fig.align='center'}
# log transform ppm
df$log_ppm <- log(df$ppm)
hist_ppm <- ggplot(df, aes(x=ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="red") 
hist_log_ppm <- ggplot(df, aes(x=log_ppm)) + 
 geom_histogram(aes(y=..density..), colour="black", fill="white", bins = 50)+
 geom_density(alpha=.2, fill="blue") 
grid.arrange(hist_ppm, hist_log_ppm, ncol=2)
```

In order to cut off extreme price values that may skew the results for typical prices, we restrict the `log_ppm` distribution to more realistic and representative values:

```{r}
df <- df %>% filter(log_ppm > -2 & log_ppm < 2)
```

## Exploratory data analysis for random intercepts

We now take a closer look at our grouping variables with a focus on the question of whether random intercepts might be appropriate.

### State

We show boxplots of `log_ppm` by `state`, arranged in the order of increasing sample size from the bottom to the top.

```{r, fig.align='center', fig.height=6}
ggplot(df, aes(x=log_ppm, y=reorder(state, state, length), fill=state)) +
geom_boxplot() +
labs(title="State", x="Log PPM by state",y="Log PPM")
```

We see that some states with small sample sizes have distributions of `log_price` that seem unstable, with respect to how much they differ from other boxplots; this provides support for the idea that the stabilizing effect of a hierarchical model would benefit our analysis. The fact that some distributions differ from each other in a noteworthy way (e.g., boxplots hardly overlapping) suggests group-based differences in price that should be taken into account in modeling, as opposed to lumping all states' data together without distinction.

### Region

Viewing a similar plot for the regions of the USA, we do see some variability in price among regions, but it is not as pronounced as for the states.

```{r, fig.align='center'}
ggplot(df, aes(x=reorder(USA_region, USA_region, length), y=log_ppm, fill=USA_region)) +
geom_boxplot() +
labs(title="log PPM by region", x="region",y="Log PPM")
```

### City

We find that there are many cities with a sample size of one, which makes `city` perhaps not the most natural choice for a grouping variable. Example visualizations shown here are for all cities in the states of Florida and California.

```{r, fig.align='center'}
fl_lppmbycity <- ggplot(df[df$state=="Florida",], aes(x=log_ppm, y=reorder(city, city, length))) +
geom_boxplot() +
labs(title="log PPM by city (FL)", x="Log PPM",y="city")
ca_lppmbycity <- ggplot(df[df$state=="California",], aes(x=reorder(city, city, length), y=log_ppm)) +
geom_boxplot() +
labs(title="log PPM by city (CA)", x="Log PPM",y="city")
grid.arrange(fl_lppmbycity, ca_lppmbycity, ncol= 2)
```

We decided to focus on `state` as our primary candidate for a grouping variable moving forward.

## Exploratory data analysis for random slopes

To consider whether we may want to include random slopes by state for our predictor variables, we examined plots of the predictor versus response among each state individually to determine whether it seemed that slopes fit to the group-level data would differ. (We present a sample of visualizations here rather than visualizations for every state, due to space constraints.)

For out continuous variables (continuous date and dosage strength), in by-state plots we often saw strikingly similar patterns in predictor vs. response, suggesting that random slopes may not be necessary for these variables.

### dose_per_mg
```{r, fig.align='center'}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = dose_per_mg, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

### date_num
```{r, fig.align='center'}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_num, y = log_ppm)) + 
  geom_point() +
  facet_grid(~ state)
```

For categorical variables, we generally observed was there was indeed variation among the states, but it was not clear whether this variation had any well-defined structure. It was not clear whether our models would benefit from random slopes for these variables, so we decided to further investigate this in our modeling process.

### bulk_purchase
```{r, fig.align='center'}
# categorical variables -- looking at South only
ggplot(data = df[df$USA_region == "South",], aes(x = bulk_purchase, y = log_ppm)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

### date_quarter
```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = date_quarter, y = log_ppm, group = date_quarter)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

### Primary_Reason
```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = Primary_Reason, y = log_ppm, group = Primary_Reason)) + 
  geom_boxplot() +
  facet_grid(~ state)
```

### source
```{r}
ggplot(data = df[df$state=="Florida" | df$state=="Texas" | df$state=="North Carolina",], aes(x = source, y = log_ppm, group = source)) + 
  geom_boxplot() +
  facet_grid(~ state)
```
## More EDA: slope and intercept distributions for predictors by state

To get a better sense of how state-level groups and slopes are distributed, we created the following intercepts for our coninuous variables:

### Dose strength variable

```{r}
group_int_list <- NULL 
group_slope_list <- NULL
for(statej in
  unique(df$state))
{
m <- lm(data = df[df$state==statej, ], log_ppm ~ dose_per_mg)
group_int_list <- c(group_int_list, coef(m)[1])
group_slope_list <- c(group_slope_list, coef(m)[2])
}
hist(group_int_list)
hist(group_slope_list)
```

### Continuous date variable

```{r}
group_int_list <- NULL 
group_slope_list <- NULL
for(statej in
  unique(df$state))
{
m <- lm(data = df[df$state==statej, ], log_ppm ~ date_num)
group_int_list <- c(group_int_list, coef(m)[1])
group_slope_list <- c(group_slope_list, coef(m)[2])
}
hist(group_int_list)
hist(group_slope_list)
```

## Model

Per the research question of interest, we are interested in exploring factors related to price per mg of oxymorphone. We are also specifically asked to control for location and how drug price changes with it. Cities seemed to be too noisy to use as a location predictor. On the flip side, region did not seem to be associated with a significant change in drug prices. Therefore, state provides the best granularity of accounting for location in this instance.

Linear regression models are usually a good standard approach to this, providing clear interpretations of how various predictors are associated with drug price. As seen in the EDA, however, many states have excessively small sample sizes. This produces corresponding $\hat{\beta}_{OLS}$ estimates with large variance. We would like to be able to *partially pool* information across states, letting us borrow information to help stabilize state estimates with small sample sizes.

This prompts a **mixed effects model**, where we instead model the state coefficients as having been drawn $N(0, \tau^2)$.

### Model Selection

Based off of the results of the EDA, we want to initially fit a model with the following variables:

- source
- dose_per_mg
- bulk_purchase
- Primary_Reason
- date_num
- date_quarter

In terms of the research question, we would also like to group based on location. There are 3 options for this:

- Region
- State
- City

In order to test whether keeping a variable in the model is "worth it", we may perform a likelihood ratio test for nested models. For adding $k$ fixed effects, the LRT test statistic follows a $\chi^2(k)$ distribution, and for adding $k$ random effects, the LRT test statistic follows a 50-50 mixture of $\chi^2(k)$ and $\chi^2(k + 1)$ distribution. For non-nested models, we can use BIC as a rough indicator of which model is "better". BIC balances likelihood of model with how parsimonious it is.

First, let's decide what to group on for region:

```{r}
# Fit 3 models grouping on different location specificity
mod_s <- lmer(data=df, log_ppm ~ (1 |state) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
mod_c <- lmer(data=df, log_ppm ~ (1 |city) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
mod_r <- lmer(data=df, log_ppm ~ (1 |USA_region) + date_num + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F)
```

```{r}
BIC <- sapply(c(mod_s, mod_c, mod_r), BIC)
re_results <- data.frame('Grouping' = c('State', 'City', 'Region'), 'BIC' = BIC)
kable(re_results, caption = 'BIC for Models with Different Random Intercepts')
```

State provides the best level of BIC, so we will choose this for our grouping variable.

Now let's investigate dropping various fixed effects and see results of our hypothesis tests:

```{r}
# Test date_num
drop_date_num <- anova(mod_s,
      lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
date_num_pval <- drop_date_num$`Pr(>Chisq)`[2]
# Test date_quarter
drop_quarter <- anova(lmer(data=df, log_ppm ~ (1 |state) + date_quarter + dose_per_mg + 
                    bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
quarter_pval <- drop_quarter$`Pr(>Chisq)`[2]
# Test dose_per_mg
drop_dose <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + bulk_purchase + Primary_Reason + source, REML=F))
dose_pval <- drop_dose$`Pr(>Chisq)`[2]
# Test bulk purchase
drop_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + Primary_Reason + source, REML=F))
bulk_pval <- drop_bulk$`Pr(>Chisq)`[2]
# Test primary reason
drop_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F))
reason_pval <- drop_reason$`Pr(>Chisq)`[2]
# Test source
drop_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason, REML=F))
source_pval <- drop_source$`Pr(>Chisq)`[2]
```

The outline is as follows:

* Start with date_num, date_quarter, dose_per_mg, bulk_purchase, Primary_Reason, source
* Perform a LRT between competing models with and without the date variable. Observe a p-value of `r round(date_num_pval, 3)` and fail to reject the null (simpler model). Drop `date_num` from the subsequent model tests
* Perform a LRT between competing models with and without the quarter of the sale. Observe a p-value of `r round(quarter_pval, 3)` and fail to reject the null (simpler model). Drop `date_quarter` from the subsequent model tests
* Perform a LRT between competing models with and without dose per mg. Observe a p-value of ~`r round(dose_pval, 3)` and reject the null (simpler model). Decide on model keeping `dose_per_mg`
* Perform a LRT between competing models with and without the bulk purchase predictor. Observe a p-value of ~`r round(bulk_pval, 3)` and reject the null (simpler model). Decide on model keeping `bulk_purchase`
* Perform a LRT between competing models with and without the primary reason predictor. Observe a p-value of `r round(reason_pval, 3)` and reject the null (simpler model). Decide on keeping `primary_reason` in the model
* Perform a LRT between competing models with and without the source predictor. Observe a p-value of ~`r round(source_pval, 3)` and reject the null (simpler model). Decide on keeping `source` in the model

Thus these results suggest we want to drop `date_num` and `date_quarter`, while keeping

* dose_per_mg
* bulk_purchase
* primary_reason
* source

We will now investigate random slopes in addition to random intercepts. We are having difficulties fitting a random slope for `dose_per_mg`, but we can check the random slope for bulk purchase:

```{r}
random_bulk <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
      lmer(data=df, log_ppm ~ (bulk_purchase |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
random_bulk$`Pr(>Chisq)`[2]  # Random slopes for bulk_purchase not helpful
random_reason <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
                       lmer(data=df, log_ppm ~ (Primary_Reason |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
random_reason$`Pr(>Chisq)`[2]
random_source <- anova(lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F),
                       lmer(data=df, log_ppm ~ (source |state) + dose_per_mg + bulk_purchase + Primary_Reason + source, REML=F))
random_source$`Pr(>Chisq)`[2]
print(paste('For testing between competing models with and without a random slope on '))
```

These results suggest that we should stick with a random intercept only.

#### Model Specification

$$log(y_{ij}) = \beta_{0,j} + \beta_{1} x_{ij1} + \beta_{2} x_{ij2} + \beta_{3} x_{ij3} + \beta_{4} x_{ij4} + \epsilon_{ij}$$

$$
\beta_{0,j} = \beta_0 + b_{0,j}, b_{0,j} \sim N(0,\tau^2)
$$
$$
\epsilon_{ij} \perp b_{0,j} \overset{iid}{\sim} N(0,\sigma^2)
$$

This linear model estimates the log price per miligram $log(y)$ of oxymorphone for an invidual $i$ in state $j$ ($x_{ij}$). The following predictors are defined as:

* $x_1$: Dosage strength in mg of the units purchased
* $x_2$: Bulk purchase, an indicator for whether 10+ units were purchased at once
* $x_3$: The primary reason for purchase. Factors include "blank", "left unanswered", "other or unknown", "self-treatment", "enjoyment", "resell", and "refuse to answer"
* $x_4$: Source of information (allows for reporting of transactions). Factors include "unkown", "internet", "internet pharmacy", "heard it", and "personal"

```{r}
# mod1b.inf <- influence(mod1b, maxfun=1)
# 
# infIndexPlot(mod1b.inf, vars='Cook')
# 
# # Find the most influential observations
# high_inf_idx <- order(cooks.distance(mod1b.inf), decreasing = TRUE)
# 
# # Take a look at some of the most influential observations
# df[high_inf_idx[1:25], ]
```

```{r}
#mod1 <- lmer(data=df, log_ppm ~ (1 | state) + date_num + mgstr + bulk_purchase + Primary_Reason + source)
#mod1b <- lmer(data=df, log_ppm ~ (1 | state) + year + mgstr + bulk_purchase + Primary_Reason + source)
# better_data <- df %>% filter(ppm > 0.2)
# mod1c <- lmer(data=better_data, log_ppm ~ (1 | state) + date_num + year + mgstr + bulk_purchase + Primary_Reason + source)
# qqmath(mod1c)
```

```{r}
mod1 <- lmer(data=df, log_ppm ~ (1 |state) + dose_per_mg + bulk_purchase + source, REML=F)
```

#### Model Diagnostics
```{r, fig.align='center', warning=FALSE, message=FALSE, error=FALSE}
# resqq <- plot_resqq(mod1)
# ranefci <- plot_ranef(mod1)
# grid.arrange(resqq, ranefci, ncol = 2)
```

Model diagnostics discussions here.